### 2014-11-13 tex/precise.tex
At VS in my edits

Initialization of uninitialized references in the $\rgse$ relation is
a non-deterministic choice that creates a heap either for
the initialization of the reference to $\mathit{null}$, a new instance of
the object, or to any type compatible object in the input heap (e.g.,
previously initialized by $\rgse$). In contrast, initialization in the
$\rsym$ relation creates a single summary heap containing all of the
possible initialization choices.

@@

The initialization rules, \figref{fig:symInit}, are invoked whenever
an uninitialized field in a symbolic reference is accessed. The
interaction with the solver in the definition of the rules is denoted
by $\mathbb{S}(\phi)$ where it returns true if $\phi$ is
satisfiable. The end rule is active when there is nothing to
initialize. This condition is determined by the function
$\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ which returns constraint-location pairs
where the field $\cfgnt{f}$ is uninitialized.
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
For reference $\cfgnt{r}$, this function constructs the set, $\theta$,
of constraint location pairs reachable from $\cfgnt{r}$ on the field
$\cfgnt{f}$ that are valid under the access path from $\cfgnt{r}$.

@@

In~\figref{fig:symInit}, given the unitialized set $\mathbb{UN}$ for
field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic (the same heap is always initialized in the same
way). 
%The $\mathrm{fresh}_\cfgnt{l}(\cfgnt{C})$ call returns a new location of type $C$. 
The set $\rho$ contains the set of
reference-location pairs from the input heap that are potential
alias. Minimizing over the set of references in the pre-image ensures
that any pair in $\rho$ is the reference $\cfgnt{r}_a$ that birthed
the location $\cfgnt{l}_a$ in some prior initialization (i.e., it was
$\cfgnt{r}_f$ and $\cfgnt{l}_f$ at one time). Aliasing constraint
in the new summary heap reason over these original creators; otherwise aliasing and local value set
computation breaks down.  The reference partitions with their
monotonicity make this determination possible.

There are three sets correspond to the non-deterministic choices in
GSE with a fourth that preserves the pre-initialized heap structure
(the constraints in each preserve the access path to the location):
(i) $\theta_\mathit{null}$ asserts the condition under which
$\cfgnt{l}_\mathit{null}$ is possible (any time the solver assigns
values to reference variables in such a way that $\cfgnt{r}_f =
\cfgnt{r}_\mathit{null}$ holds as does access path $\phi_x$). (ii)
$\theta_\mathit{new}$, meaning $\cfgnt{r}_f$ refers to the fresh
location, occurs when $\cfgnt{r}_f$ is not new or any other alias. (iii) any member of $\theta_\mathit{alias}$ restricts $\cfgnt{r}_f$ to
only alias one member of $\rho$ and not alias any member of $\rho$
that was initialized earlier than the current choice (a later
initialization may indeed alias the current alias choice in a valid
heap but the current choice cannot alias both the choice and a
previous initialization). And (iv) $\theta_\mathit{orig}$  implements conditional initialization to preserve the
original heap structure. This property is needed for the bisumulation. 
These sets are added into the heap on the field after initializing the fields in the new location to $\cfgnt{l}_\mathit{un}$.

@@

\begin{figure*}[t]
\begin{center}
\begin{tabular}[c]{c|c|c|c}
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{origHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeXHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeYHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{l}
$\rho := \{ (\cfgnt{r}_1^i\ \cfgnt{l}_1 \}$ \\
$\theta_\mathit{null} := \{ ( \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}\ \cfgnt{l}_\mathit{null}) \}$\\
$\theta_\mathit{new} := \{ ( \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq \cfgnt{r}_1^i\ \cfgnt{l}_2) \}$\\
$\theta_\mathit{alias} := \{ ( \cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i\ \cfgnt{l}_1) \}$\\
$\theta_\mathit{orig} := \{ \}$ \\
$\phi_{\mathit{1a}} := \cfgnt{r}_1^i = \cfgnt{r}_\mathit{null} $ \\
$\phi_{\mathit{1b}} := \cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null} $  \\
$\phi_{\mathit{2a}} := \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}$ \\
$\phi_{\mathit{2b}} := \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq \cfgnt{r}_1^i$ \\
$\phi_{\mathit{2c}} :=  \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i $ \\
\end{tabular} \\
(a) & (b) & (c) & (d) \\
\end{tabular}
\end{center}
\caption{An example that initializes $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ and $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. (a) Initial heap structure. (b) After $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ is initialized. (c) After $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ is initialized. (d) Constraint sets from the summary rule and constraints on the edges.}
\label{fig:initHeap}
\end{figure*}

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a reference from the stack partition.
 In~\figref{fig:initHeap} $\cfgnt{r}_0^s$ represents a stack
reference for the $\cfgt{this}$ variable which has two fields $x$
and $y$ of the same type; reference $\cfgnt{r}_0^s$ points to location
$\cfgnt{l}_0$. Note that when no constraint is specified in the graphs, then
there is an implicit $\mathit{true}$ constraint, for example, $\cfgnt{r}_0^s$
points to $\cfgnt{l}_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $\cfgnt{r}_\mathit{un}$.

@@

The access on $\cfgnt{l}_0.x$ creates a new input heap reference $\cfgnt{r}_1^i$ in the
summary heap shown in~\figref{fig:initHeap}(b). The reference $\cfgnt{r}_1^1$
points to the $\cfgnt{l}_\mathit{null}$ location under the constraint that
$\cfgnt{r}_1^i$ is null: $\phi_{1a} := \cfgnt{r}_1^i = \cfgnt{r}_\mathit{null}$. The reference
$\cfgnt{r}_1^i$ points to the location $\cfgnt{l}_1$ under the constraint that $\cfgnt{r}_1^i$
is not $\mathit{null}$ $\phi_{1b} :=\cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null}$. The
location $\cfgnt{l}_1$ represents a fresh location of type $C$ is created on
the heap such that $C$ is type of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$.

Following initialization to $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$, when
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ is initialized, the resulting summary heap is shown
in~\figref{fig:initHeap}(c). The set $\rho$ shown
in~\figref{fig:initHeap}(d) contains the potential aliases for
 $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ which is location $\cfgnt{l}_1$ ($\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$). The
constraint location $\theta$ sets represents the initialization
choices for $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. The new input reference $\cfgnt{r}_2^i$ for
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ points to (a) location $\cfgnt{l}_\mathit{null}$ under the constraint
$\phi_{2a} := \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}$, (b) location $\cfgnt{l}_2$ under the
constraint $\phi_{2b} := \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq
\cfgnt{r}_1^i$ that states that $\cfgnt{r}_2^i$ is not null and does not alias
$\cfgnt{r}_1^i$, and (c) location $\cfgnt{l}_1$ under the constraint $\phi_{2c} :=
\cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i$ that states $\cfgnt{r}_2^i$
is not null and \emph{does} alias $\cfgnt{r}_1^i$.

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). In both the field rewrite rules we first check we
perform is whether there exists a constraint location pair for the $r$
being accessed such that the location is not null and the constraint
when conjuncted with the global constraint is satisfiable. Next we
extract all possible constraints under which $r$ points to a null
location such that the constraint is satisfiable under the current
global constraint, $\phi_g$. The negation of these constraints are
added to the global constraint to create a new global constraint
$\phi_g^\prime$. The update to the global constraint ensures that
access or write of the field $f$ happens only on non-null
locations. There are rewrite rules for accessing and writing to fields
on null locations where it leads the current program execution state
to an error state. The rewrite rules for the accessing and writing to
null locations are presented in the supplemental document.

In the field-access rewrite rule in~\figref{fig:fHeap}(a) after the
non-nullness check we extract the The type $C$ of the field is
extracted and invoke the summarize rewrite rule
in~\figref{fig:symInit} that performs the initialization of the field
if it points to uninitialized locations. Once the initialization is
complete, we create a new local reference $r^\prime$. An important
property of the references in the bi-partiate graph is that they are
\emph{immutable}. Hence we create a new local reference and assign it
to the initialized reference, and return the local reference. In order
to de-reference a field $r.f$ we define a helper function which is
called the value set.

>>>>

Initialization of uninitialized references in the $\rgse$ relation is
a non-deterministic choice that creates a heap either for
the initialization of the reference to null, a new instance of
the object, or to any type compatible object in the input heap (e.g.,
previously initialized by $\rgse$). In contrast, initialization in the
$\rsym$ relation creates a single summary heap containing all of the
possible initialization choices.

@@

The initialization rules, \figref{fig:symInit}, are invoked whenever
an uninitialized field in a symbolic reference is accessed. The
interaction with the solver in the definition of the rules is denoted
by $\mathbb{S}(\phi)$ where it returns true if $\phi$ is
satisfiable. The end rule is active when there is nothing to
initialize. This condition is determined by the function
$\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ which returns constraint-location pairs
where the field $\cfgnt{f}$ is uninitialized.
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
A pair in the set means the field in the location maps to a reference
that then points to $\cfgnt{l}_\mathit{un}$, and that the path to that
location is feasible on some aliasing assignment.

@@

In~\figref{fig:symInit}, given the unitialized set $\Lambda$ for
field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic (i.e., the same heap is always initialized in the same
way as order determines how many aliases are available in the input heap at any given point in initialization). 
%The $\mathrm{fresh}_\cfgnt{l}(\cfgnt{C})$ call returns a new location of type $C$. 
The set $\rho$ contains the set of
reference-location pairs from the input heap that are potential
aliases. Minimizing over the set of references in the pre-image ensures
that any pair in $\rho$ is the reference $\cfgnt{r}_a$ and the location $\cfgnt{l}_a$ materialized at the same time in some prior initialization (i.e., they were once
$\cfgnt{r}_f$ and $\cfgnt{l}_f$). Aliasing constraints
in the new summary heap reason over these original creators; otherwise aliasing and local value set
computation breaks down.  The reference partitions with their
monotonicity make this determination possible.

There are three sets corresponding to the non-deterministic choices in
GSE with a fourth that preserves the pre-initialized heap structure.
Note that the constraints always preserve the access path to the location which is $\phi_x$.
(i) $\theta_\mathit{null}$ asserts the condition under which
$\cfgnt{l}_\mathit{null}$ is possible (any time the solver assigns
values to reference variables in such a way that both $\phi_x$ and $\cfgnt{r}_f =
\cfgnt{r}_\mathit{null}$ hold). (ii)
$\theta_\mathit{new}$ is active when $\cfgnt{r}_f$ points to the
fresh location. This occurs when $\cfgnt{r}_f$ is not new or any other alias. (iii) a member of $\theta_\mathit{alias}$ restricts $\cfgnt{r}_f$ to
a particular alias in $\rho$, and at the same time, not alias any member of $\rho$
that was initialized earlier than the current choice (a later
initialization may indeed also alias the current choice in a valid
heap as order determines the number of aliasing choices available). And (iv) $\theta_\mathit{orig}$  implements conditional initialization to preserve the
original heap structure. This property is needed for the bisumulation. 
These sets are added into the heap on $\cfgnt{r}_f$ after the fields for $\cfgnt{l}_f$ are initialized to $\cfgnt{r}_\mathit{un}$.

@@

\begin{figure*}[t]
\begin{center}
\begin{tabular}[c]{c|c|c|c}
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{origHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeXHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{c}
\scalebox{0.81}{\input{summarizeYHeap.pdf_t}} \\
\end{tabular} &
\begin{tabular}[c]{l}
$\rho := \{ (\cfgnt{r}_1^i\ \cfgnt{l}_1 \}$ \\
$\theta_\mathit{null} := \{ ( \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}\ \cfgnt{l}_\mathit{null}) \}$\\
$\theta_\mathit{new} := \{ ( \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq \cfgnt{r}_1^i\ \cfgnt{l}_2) \}$\\
$\theta_\mathit{alias} := \{ ( \cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i\ \cfgnt{l}_1) \}$\\
$\theta_\mathit{orig} := \{ \}$ \\
$\phi_{\mathit{1a}} := \cfgnt{r}_1^i = \cfgnt{r}_\mathit{null} $ \\
$\phi_{\mathit{1b}} := \cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null} $  \\
$\phi_{\mathit{2a}} := \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}$ \\
$\phi_{\mathit{2b}} := \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq \cfgnt{r}_1^i$ \\
$\phi_{\mathit{2c}} :=  \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i $ \\
\end{tabular} \\
(a) & (b) & (c) & (d) \\
\end{tabular}
\end{center}
\caption{An example that initializes $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ and $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. (a) Initial heap structure. (b) After $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ is initialized. (c) After $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ is initialized. (d) Sets in the summarize rule and constraints on the edges.}
\label{fig:initHeap}
\end{figure*}

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a reference from the stack partition.
 In~\figref{fig:initHeap}, $\cfgnt{r}_0^s$ represents a stack
reference for the $\cfgt{this}$ variable which has two fields $x$
and $y$ of the same type. Note that when no constraint is specified in the graphs, then
there is an implicit $\mathit{true}$ constraint, for example, $\cfgnt{r}_0^s$
points to $\cfgnt{l}_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $\cfgnt{r}_\mathit{un}$.

@@

The access of $x$ from $\cfgnt{l}_0$ creates a new input heap reference $\cfgnt{r}_1^i$ in the
summary heap shown in~\figref{fig:initHeap}(b). The reference $\cfgnt{r}_1^i$
points to the $\cfgnt{l}_\mathit{null}$ location under the constraint that
$\cfgnt{r}_1^i$ is null (\figref{fig:inintHeap}(d)): $\phi_{1a} := \cfgnt{r}_1^i = \cfgnt{r}_\mathit{null}$. The reference
$\cfgnt{r}_1^i$ points to the location $\cfgnt{l}_1$ under the constraint that $\cfgnt{r}_1^i$
is not $\mathit{null}$ ($\phi_{1b}$). The
location $\cfgnt{l}_1$, a fresh location of type $C$ (the type of field $\cfgnt{x}\rp$), is added to the heap. There are no aliases because the input heap was empty.

Following $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$, 
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ is initialized leading to the summary heap
in~\figref{fig:initHeap}(c). The set $\rho$ 
in~\figref{fig:initHeap}(d) contains the one potential aliases $\cfgnt{l}_1$ from the previous initialization for $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$. It also shows the $\theta$-sets for this initialization and the resulting edge labels. Note that the ordering between aliases in $\theta_\mathit{alias}$ does not matter until a third field is initialized and there are two potential choices.

\begin{comment}
The
constraint location $\theta$ sets represents the initialization
choices for $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. The new input reference $\cfgnt{r}_2^i$ for
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ points to (a) location $\cfgnt{l}_\mathit{null}$ under the constraint
$\phi_{2a} := \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}$, (b) location $\cfgnt{l}_2$ under the
constraint $\phi_{2b} := \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq
\cfgnt{r}_1^i$ that states that $\cfgnt{r}_2^i$ is not null and does not alias
$\cfgnt{r}_1^i$, and (c) location $\cfgnt{l}_1$ under the constraint $\phi_{2c} :=
\cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i$ that states $\cfgnt{r}_2^i$
is not null and \emph{does} alias $\cfgnt{r}_1^i$.
\end{comment}

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location. If such is the case, then the global
constraint $\phi_g$ is updated to disallow any viable null locations
from this point forward: future accesses are non-null.
%perform is whether there exists a constraint location pair for the $r$
%being accessed such that the location is not null and the constraint
%when conjuncted with the global constraint is satisfiable. Next we
%extract all possible constraints under which $r$ points to a null
%location such that the constraint is satisfiable under the current
%global constraint, $\phi_g$. The negation of these constraints are
%added to the global constraint to create a new global constraint
%$\phi_g^\prime$.
%The update to the global constraint ensures that
%access or write of the field $f$ happens only on non-null
%locations. 
When there is a feasible null location in the summary heap on the
field access operations, the machine moves to an error state. The
rewrite rules for null locations are in the supplemental document.

In the field-access rewrite rule in~\figref{fig:fHeap}(a), after the
non-nullness check we extract the The type $C$ of the field is
extracted and invoke the summarize rewrite rule
in~\figref{fig:symInit} that performs the initialization of the field
if it points to uninitialized locations. Once the initialization is
complete, we create a new local reference $r^\prime$. An important
property of the references in the bi-partiate graph is that they are
\emph{immutable}. Hence we create a new local reference and assign it
to the initialized reference, and return the local reference. In order
to de-reference a field $r.f$ we define a helper function which is
called the value set.
